Learning Auditory Similarities Between Instrumental Playing Techniques

### 2019-09-26 paper/paper.tex
Fix name and affiliation for JA
Write discussion section

\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   noteref={n1},                        % id's of article notes, if any
   email={email@example.com}            % email address
]{\inits{VL}\fnm{Vincent} \snm{Lostanlen}}
\author[
   addressref={aff2},
   email={email@example.com}
]{\inits{CE}\fnm{Christian} \snm{El-Hajj}}
\author[
   addressref={aff2},
   email={email@example.com}
]{\inits{MR}\fnm{Mathias} \snm{Rossignol}}
\author[
   addressref={aff2},
   email={email@example.com}
]{\inits{GL}\fnm{Gr\'egoire} \snm{Lafay}}
\author[
   addressref={aff2},
   email={email@example.com}
]{\inits{JA}\fnm{Joakim} \snm{And\`en}}
\author[
   addressref={aff2},
   email={email@example.com}
]{\inits{ML}\fnm{Mathieu} \snm{Lagrange}}

@@

\address[id=aff1]{%                           % unique id
  \orgname{Department of Zoology, Cambridge}, % university, etc
  \street{Waterloo Road},                     %
  %\postcode{}                                % post or zip code
  \city{London},                              % city
  \cny{UK}                                    % country
}
\address[id=aff2]{%
  \orgname{Marine Ecology Department, Institute of Marine Sciences Kiel},
  \street{D\"{u}sternbrooker Weg 20},
  \postcode{24105}
  \city{Kiel},
  \cny{Germany}
}

@@

\begin{itemize}
\item Demonstrates complex timbral structure perceived by subjects in study: neither instrument (i.e., spectral envelope) or playing technique (i.e., temporal modulation) dominates similarity judgment. Need joint to characterize time-frequency structure (see Patil et al).
\item Proposed computational model (scattering + linear) reproduces similarity judgments accurately. Easy to ``retrain'' for other data. Fixed wavelet structure allows for analysis of linear layer and guarantees invariance and stability properties.
\item Applications: query-by-example in musical sample databases.
\end{itemize}

>>>>

\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   noteref={n1},                        % id's of article notes, if any
   email={email@example.com}            % email address
]{\inits{VL}\fnm{Vincent} \snm{Lostanlen}}
\author[
   addressref={aff2},
   email={email@example.com}
]{\inits{CE}\fnm{Christian} \snm{El-Hajj}}
\author[
   addressref={aff2},
   email={email@example.com}
]{\inits{MR}\fnm{Mathias} \snm{Rossignol}}
\author[
   addressref={aff2},
   email={email@example.com}
]{\inits{GL}\fnm{Gr\'egoire} \snm{Lafay}}
\author[
   addressref={ccm},
   email={janden@flatironinstitute.org}
]{\inits{JA}\fnm{Joakim} \snm{And\'en}}
\author[
   addressref={aff2},
   email={email@example.com}
]{\inits{ML}\fnm{Mathieu} \snm{Lagrange}}

@@

\address[id=aff1]{%                           % unique id
  \orgname{Department of Zoology, Cambridge}, % university, etc
  \street{Waterloo Road},                     %
  %\postcode{}                                % post or zip code
  \city{London},                              % city
  \cny{UK}                                    % country
}
\address[id=aff2]{%
  \orgname{Marine Ecology Department, Institute of Marine Sciences Kiel},
  \street{D\"{u}sternbrooker Weg 20},
  \postcode{24105}
  \city{Kiel},
  \cny{Germany}
}
\address[id=ccm]{%
  \orgname{Center for Computational Mathematics, Flatiron Institute},
  \street{162 5th Avenue},
  \city{New York, NY 10010},
  \cny{USA}
}

@@

The success of the above system in identifying timbral similarities has immediate applications in browsing music databases.
These are typically organized based on instrumental and PT taxonomies, with additional keywords offering a more flexible organization.
Accessing the sounds in these databases therefore requires some knowledge of the taxonomy and keywords used.
Furthermore, the user needs to have some idea of the particular IPT they are searching for.

As an alternative, content-based searches allow the user to identify sounds based on similarity to some given query sound.
This query-by-example approach provides an opportunity to search for sounds without having a specific instrument of PT in mind, yielding a wider range of available sounds.
A composer with access to such a system would therefore be able to draw on a more diverse palette of musical timbre.

The computational model proposed in this work is well suited to such a query-by-example task.
We have shown that it is able to approximate well the timbral judgments of a wide range of subjects included in our study.
Not only that, but the system can be easily retrained to approximate an individual user's timbre perception by having that user perform the clustering task on a reduced set of $78$ IPTs and running the LMNN training step on those clustering assignments.
This model can then be applied to new data, or, alternatively, be retrained with these new examples if the existing model proves unsatisfactory.

We shall note, however, that the current model has several drawbacks.
First, it is only applied to instrumental sounds.
While this has the advantage of simplifying the intepretation of the results, the range of timbre under consideration is necessarily limited (although less restricted than only considering \emph{ordinario} PTs).
This also makes applications such as \emph{query-by-humming} difficult, since we cannot guarantee that the timbral similarity measure is accurate for such sounds.

That being said, the above model is general enough to encompass a wide variety of recordings, not just instrumental sounds.
Indeed, the tools used (scattering transforms and LMNN weighting matrices) do not depend on the type of sound being processed and work for more general classes of sounds.
To extend the model, it is only necessary to retrain the LMNN weighting matrix by supplying it with new cluster assignments.
These can again be obtained by performing a new clustering experiment with one or more human subjects.

Another aspect is the granularity of the similarity judgments.
In the above method, we have used \emph{hard} clustering assignments to build our model.
A more nuanced similarity judgment would ask users to rate the similarity of a pair of IPTs on a more graduated scale, which would yield a finer, or \emph{soft}, assignment.
This however, comes with additional difficulties in providing a consistent scale across subjects, but could be feasible if the goal is to only adapt the timbral space to a single individual.
An approach not based on clustering would also have to replace the LMNN algorithm with one that accounts for such soft assigments.
